Fast Algorithms for Intersection of Non-matching Grids Using \plucker Coordinates.

### 2016-10-01 paper_intersections.tex
Improve prolongation scheme, delete pdf.
Improve advancing front algorithm.

@@

\section{Global mesh intersection algorithm}
\label{sec:front_advancing}
Having the algorithms for element-element intersections we can proceed to the mesh intersection algorithm. 
We first compute all component-bulk mesh intersections that is (1d-3d and 2d-3d) and then the 1d-2d and 2d-2d intersections use the 3d mesh
to get the intersection pair candidates.  Element intersections for the particular component and bulk mesh 
are obtained in two phases: firstly, we look for the first two elements intersecting each other (initialization); 
secondly, we prolong the intersection by investigating neighbouring elements (front tracking).
\noteJB{Not sure that "prolong" is the right verb.}

@@

\paragraph{Initialization}
We choose an arbitrary element of the component mesh and check it intersection with all elements of the bulk mesh.
The process is accelerated by computing the axis aligned bounding box (AABB) for every element and use intersection of the bounding boxes
as a fast indicator for possible intersection of the elements. This step takes $O(N)$ time with respect to the number of elements of the bulk mesh $N$.
For the small number of components $k$ and if they are contained inside the bulk mesh the total time of initialization may be still linear $O(kn)$. However,
for more complex cases we organize the bounding boxes of the bulk mesh into the bounding interval hierarchy (BIH) \cite{EGWR:EGSR06:139-149} 
a data structure in principle equivalent 
to R-trees \cite{guttman_r-trees:_1984}, \cite{nam_comparative_2004}. The creation time is $O(N\log(N/n))$ and the lookup time $O(\log(N/n))$ where $n$ is number
of the bulk elements in the leaf nodes of the tree.

\paragraph{Advancing Front Method}
The advancing front algorithm requires and employ the neigbouring information for the elements within the component as well as within the bulk mesh.
It can be viewed as the breadth first search algorithm for a graph which vertices are intersection polygons and edges are sides shared by
two polygons. Side every side of IP is on the boundary of either component or the bulk element we can distinguish bulk and component edges and correspondingly 
the \emph{bulk queue} and the \emph{component queue}.

%
\begin{figure}[!htb]
%   \vspace{0pt}
  \centering    
    \includegraphics[width=\textwidth]{\figpath prolongation_scheme_ip.pdf}
  \caption{Advancing front algorithm for 1d-2d and 2d-3d intersections.}
  \label{fig:prolongation}
\end{figure}

The scheme of the advancing front algorithm is displayed by the scheme in \fig{fig:prolongation}. 
The main idea is to compute intersections for a component element with all possible bulk elements, and then move to a~next neighboring component element. 
So the bulk queue is emptied before the component queue (see yellow boxes).

On the input we consider a candidate pair provided be the initialization (top-left corner).
Now we look for new candidate pairs among the neighboring elements (the block Prolongation decide).
Therefore, we iterate over the intersection corners and further exploit their topological position on the component element $c$ and the bulk element $b$.
There are 3 possible cases (applies both for 1d-3d and 2d-3d), how the intersection might be prolonged:
\begin{itemize}
    \item \textbf{IC on the boundary of $c$  and inside $b$} \\
            We find all the sides $S$ coincident with the \nface of $c$ on which IC lies. Then we get all component
            elements $C'$ neighbouring with $c$ over any side$s\in S$. And finally, we push all pairs $(c',b)$, $c'\in C'$ 
            into the component queue.
            Note, that $c$ can have more then one neighbour component elements over the single side $s$, i.e. branches are allowed.
    \item \textbf{IC on the boundary of both $c$ and $b$} \\
            We find all the faces $F$ coincident with the \nface of $b$ on which IC lies. 
            Then we get all neighbours $B'$ through all faces in $F$ as in the previous case.
            We push pairs $(c, b')$, $b'\in B'$ into the bulk queue and pairs $(c', b')$, $c'\in C'$, $b'\in B'$ into the component queue.
    \item \textbf{IC inside of $c$ anf on the boundary of $b$)} \\
            We proceed as in previous case, and push pairs $(c, b')$, $b'\in B'$ into the bulk queue.
\end{itemize}
\noteJB{Can we just use two (not excluded) cases one for IB on boundary of $c$ second for CI on boundary of $b$ ?}

We also see that the candidate pairs are of three types: $(c, b')$, $(c', b)$, $(c',b')$
from which only the first one goes into the bulk queue, trying to cover current component element $c$ before we proceed to the next one.

When the pair is taken from either of the queues we check if the pair was computed and possibly skip it. 
To this end every component element keeps a hash of the bulk elements with which we have called the intersection test.

>>>>

% alphabetical enumeration
\usepackage{enumitem}

@@

\section{Global Mesh Intersection Algorithm}
\label{sec:front_advancing}
Having the algorithms for element-element intersections we can proceed to the mesh intersection algorithm. 
We first compute all component-bulk mesh intersections that is (1d-3d and 2d-3d) and then the 1d-2d and 2d-2d intersections using the 3d mesh to get the intersection pair candidates.
Element intersections for the particular component and bulk mesh 
are obtained in two phases: firstly, we look for the first two elements intersecting each other (initialization); 
secondly, we prolong the intersection by investigating neighbouring elements (front tracking).
\noteJB{Not sure that "prolong" is the right verb.}

@@

\paragraph{Initialization}
We choose an arbitrary element of the component mesh and check its intersection with all elements of the bulk mesh.
This process is accelerated by computing the axis aligned bounding box (AABB) for every element and use intersection of the bounding boxes
as a fast indicator for possible intersection of the elements. This step takes $O(N)$ time with respect to the number of elements of the bulk mesh $N$.
If the number of components $k$ is small and if the components are contained inside the bulk mesh, the total time of initialization may still be linear $O(kn)$ \notePE{I believe you ment $O(kN)$ here}. However,
for more complex cases we organize the bounding boxes of the bulk mesh into the bounding interval hierarchy (BIH) \cite{EGWR:EGSR06:139-149},
a data structure in principle equivalent 
to R-trees \cite{guttman_r-trees:_1984}, \cite{nam_comparative_2004}. The creation time is $O(N\log(N/n))$ and the lookup time $O(\log(N/n))$ where $n$ is number
of the bulk elements in the leaf nodes of the tree.

\paragraph{Advancing Front Method}
The advancing front method requires the neigbouring information for the elements within the components as well as within the bulk mesh.
It can be viewed as the breadth first search algorithm for a graph which vertices are intersection polygons and edges are sides shared by
two polygons. Since every side of IP is on the boundary of either component or bulk element, we can distinguish the bulk and component edges. Correspondingly, we save new candidate pairs into a~\emph{bulk queue} or 
a~\emph{component queue} to be subsequently processed later.
The scheme of the advancing front algorithm is displayed by the scheme in \fig{fig:prolongation},
where the queues are represented by the yellow boxes.
%
\begin{figure}[!htb]
%   \vspace{0pt}
  \centering    
    \includegraphics[width=\textwidth]{\figpath prolongation_scheme_ip_new.pdf}
  \caption{Advancing front algorithm for 1d-2d and 2d-3d intersections.}
  \label{fig:prolongation}
\end{figure}

The key idea behind the two queues is to compute intersections for a component element with all possible bulk elements at first, and then move to a~next neighboring component element. So the bulk queue is emptied before the component queue.

On the input, we consider a candidate pair provided by the initialization (the top-left corner of the scheme).
If a nonempty intersection is found, we look for new candidate pairs among the neighboring elements
(see the white block 'Cycle over ICs').
Therefore, we iterate over the intersection corners and further exploit their topological position on the component element $c$ and the bulk element $b$.
There are 3 possible cases (applies both for 1d-3d and 2d-3d), how the intersection might be prolonged:
\begin{enumerate}[label=(\alph*)]
    \item \textbf{IC on the boundary of $c$  and inside $b$} \\
            We find all the sides $S$ coincident with the \nface of $c$ on which IC lies. Then we get all component
            elements $C'$ neighboring with $c$ over any side $s\in S$. And finally, we push all pairs $(c',b)$, $c'\in C'$ 
            into the component queue.
            Note that $c$ can have more then one neighbour component elements over the single side $s$, i.e. branches are allowed.
    \item \textbf{IC inside of $c$ anf on the boundary of $b$} \label{enum:prolong2}\\
            We find all the faces $F$ coincident with the \nface of $b$ on which IC lies. 
            Then we get all bulk elements $B'$ neighboring with $b$ over any face $f\in F$,
            analogically to the previous case.
            Finally, we push the new candidate pairs $(c, b')$, $b'\in B'$ into the bulk queue.
    \item \textbf{IC on the boundary of both $c$ and $b$} \\
            This situation includes both previously described cases. Therefore, we obtain new candidate pairs
            $(c', b)$, $c'\in C'$ and $(c, b')$, $b'\in B'$ and push them into the component queue and the bulk queue, respectively.
\end{enumerate}
% \noteJB{Can we just use two (not excluded) cases one for IB on boundary of $c$ second for CI on boundary of $b$ ?}
When the pair is taken from either of the queues, we check whether the pair has been already computed and possibly skip it. 
To this end every component element keeps a hash of the bulk elements with which we have called the intersection test.

At the end, we see that the candidate pairs are of two types: $(c, b')$ and $(c', b)$,
from which the first one goes into the bulk queue, trying to fully cover current component element 
$c$ before we proceed to the next one. We are able recognize the case when the component element 
does not lie inside the bulk elements with its full volume -- we end up in situation \ref{enum:prolong2},
but there are no more bulk neighboring elements to prolong the intersection to. Then we mark this component
element as \emph{open}, so it can be later found by initialization algorithm and possibly
covered by other bulk elements but as different component.
